% -*- coding:utf-8; mode:latex; -*- %

% Copyright (C) 2018 Moritz Klammler <moritz.klammler@student.kit.edu>
%
% This work is licensed under a Creative Commons Attribution-NonCommercial-NoDerivatives 4.0 International License
% (https://creativecommons.org/licenses/by-nc-nd/4.0/).

\DeclareAcronym{chap}{%
  short = \textsection,
  short-plural = \textsection,
  long = chapter,
  sort = {*},
  class = internal,
}

\DeclareAcronym{section}{%
  short = \textsection,
  short-plural = \textsection,
  long = section,
  sort = {*},
}

\DeclareAcronym{paragraph}{%
  short = \textparagraph,
  short-plural = \textparagraph,
  long = paragraph,
  sort = {*},
  class = internal,
}

\DeclareAcronym{ad}{%
  short = {a.d.},
  long = {also known as},
  foreign = {alias dictus},
  short-format = \itshape,
  foreign-lang = latin,
}

\DeclareAcronym{algo}{%
  short = {algo.},
  long = algorithm,
}

\DeclareAcronym{ascii}{%
  short = {ASCII},
  long = {American Standard Code for Information Interchange},
}

\DeclareAcronym{cf}{%
  short = {cf.},
  long = {compare, consult},
  foreign = {confer},
  short-format = \itshape,
  foreign-lang = latin,
}

\DeclareAcronym{def}{%
  short = {def.},
  long = {definition},
}

\DeclareAcronym{ed}{%
  short = {ed.},
  long = {edition, editor, edited},
}

\DeclareAcronym{eg}{%
  short = {e.g.},
  long = {for example},
  foreign = {exempli gratia},
  short-format = \itshape,
  foreign-lang = latin,
}

\DeclareAcronym{eq}{%
  short = {eq.},
  long = {equation},
}

\DeclareAcronym{etal}{%
  short = {et~al.},
  long = {and others},
  foreign = {et alii},
  short-format = \itshape,
  foreign-lang = latin,
}

\DeclareAcronym{fig}{%
  short = {fig.},
  long = {figure},
}

\DeclareAcronym{fmmm}{%
  short = FMMM,
  alt = {FM\textsuperscript{3}},
  long = {Fast Multipole Multi-Level Method},
}

\DeclareAcronym{graphml}{%
  short = GraphML,
  long = {Graph Markup Language},
}

\DeclareAcronym{html}{%
  short = HTML,
  long = {Hypertext Markup Language},
}

\DeclareAcronym{id}{%
  short = ID,
  long = {identity, unique identifier},
}

\DeclareAcronym{ie}{%
  short = {i.e.},
  long = {that is},
  foreign = {id est},
  short-format = \itshape,
  foreign-lang = latin,
}

\DeclareAcronym{io}{%
  short = {I/O},
  long = {input {\&} output},
}

\DeclareAcronym{json}{%
  short = JSON,
  long = {JavaScript Object Notation},
}

\DeclareAcronym{mse}{%
  short = {MSE},
  long = {mean squared error},
}

\DeclareAcronym{nist}{%
  short = NIST,
  long = {National Institute of Standards and Technology},
}

\DeclareAcronym{ogdf}{%
  short = OGDF,
  long = {Open Graph Drawing Framework},
}

\DeclareAcronym{pc}{%
  short = {PC},
  long = {personal computer},
}

\DeclareAcronym{pca}{%
  short = PCA,
  long = {principal component analysis},
}

\DeclareAcronym{posix}{%
  short = POSIX,
  long = {Portable Operating System Interface},
}

\DeclareAcronym{pp}{%
  short = {pp.},
  long = {pages},
}

\DeclareAcronym{ram}{%
  short = {RAM},
  long = {random access memory},
}

\DeclareAcronym{rdf}{%
  short = RDF,
  long = {radial distribution function},
}

\DeclareAcronym{relu}{%
  short = {ReLU},
  long = {rectified linear unit},
}

\DeclareAcronym{rms}{%
  short = RMS,
  long = {root mean squared},
}

\DeclareAcronym{sgd}{%
  short = {SGD},
  long = {stochastic gradient descent},
}

\DeclareAcronym{sql}{%
  short = {SQL},
  long = {Structured Query Language},
}

\DeclareAcronym{st}{%
  short = {s.t.},
  long = {such that},
}

\DeclareAcronym{tab}{%
  short = {tab.},
  long = {table},
}

\DeclareAcronym{tar}{%
  short = {TAR},
  long = {tape archive},
}

\DeclareAcronym{thm}{%
  short = {thm.},
  long = {theorem},
}

\DeclareAcronym{ui}{%
  short = {UI},
  long = {user interface},
}

\DeclareAcronym{url}{%
  short = URL,
  long = {uniform resource locator},
}

\DeclareAcronym{wrt}{%
  short = {w.r.t.},
  long = {with respect / regard to},
}

% If the following acronym is renamed to 'xml' (as it ought to) the formatting of the list of abbreviations breaks.
% I frankly cannot understand why this happens but it took me the better part of a day to figure out.
\DeclareAcronym{nxml}{%
  short = XML,
  long = {Extensible Markup Language},
  sort = {xml},
}

\DeclareAcronym{nxslt}{%
  short = XSLT,
  long = {Extensible Stylesheet Language Transformations},
  sort = {xslt},
}

\newcommand*{\fakeacronyms}{%
  \acuse{pp,ed,etal}
}
